\textls[-0]{\thetitle}

### 2023-03-21 metatheory2.tex
Working on cover letter.

\Citeauthor{debruijn1964} indices and explicit substitutions make details about
binders and substitutions explicit.
%The change of the syntax formulation is important to the simplicity of our
%proof, as we want to handle the proof about substitutions and terms
%uniformly---thus making substitution as part of the syntax (explicit
%substitution) will be helpful. 
\EDJ{Why comment out this sentence?}
The form \var{n} represents a variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Substitutions are typed with the form 
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
%
The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
where $\SubstId$ is the identity substitution (\ruleref{sub/id}).

@@

>>>>

\Citeauthor{debruijn1964} indices and explicit substitutions make details about
binders and substitutions explicit.
%
%The change of the syntax formulation is important to the simplicity of our
%proof, as we want to handle the proof about substitutions and terms
%uniformly---thus making substitution as part of the syntax (explicit
%substitution) will be helpful. 
\EDJ{Why comment out this sentence?}
%
The form \var{n} represents a variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Substitutions are typed with the form 
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
%
The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
where $\SubstId$ is the identity substitution (\ruleref{sub/id}).

@@

Consequently, function application changes to a equivalent formulation, and becomes a ``direct inverse'' of typing rule for function abstraction. For example, the named notation $\napp{f}{t}$ can be equivalently represented by $\sub{\app{f}}{\SubstExt{\SubstId}{t}}$.
